---
title: "LEGO-spreadsheet-QA"
author: "Rick O. Gilmore"
date: "`r Sys.time()`"
output: 
  html_document:
    toc: true
    toc_levels: 3
    toc_float: true
    number_sections: true
    code_folding: hide
params: 
  db_account: rogilmore@psu.edu
---

# Purpose

This document summarizes the quality assurance (QA) process conducted on the LEGO project session/participant metadata spreadsheet.
The data are stored and shared at <https://nyu.databrary.org/volume/563>.

# Set-up

```{r setup, include=TRUE}
knitr::opts_chunk$set(echo = TRUE)

# Load databraryapi package from github.com/PLAY-behaviorome/databraryapi
if (!require(databraryapi)) {
  devtools::install_github("PLAY-behaviorome/databraryapi")
  library(databraryapi)
}

# Load other libraries
library(tidyverse)
library(ggplot2)

## Variables for this particular volume and session.
play_vol_id <- 563
#survey_data_session_name = "Parent report data"
```

The Databrary volume is 563.

We need to log in to Databrary to have access to these data.
If you have logged-in before from the machine you are running this report, then you will be prompted to enter your Databrary ID (email address).
If you have *not* logged-in before, you will be prompted to enter your Databrary ID (email address) and your Databrary password.

You may wish to knit (update) this report from the RStudio command line using this command: `rmarkdown::render('LEGO-spreadsheet.Rmd', params=list(db_account="youremail@yourdomain.com"))`, where `youremail@yourdomain.com` is your Databrary account ID.

```{r login}
databraryapi::login_db(params$db_account)
```

# Gather data

Participant demographic data are stored in the Databrary spreadsheet.

## Participant demographics

We'll first gather the participant demographic data stored in the Databrary spreadsheet.

```{r gather-participant-demographic-data}
lego_demog_raw <- databraryapi::download_session_csv(vol_id = play_vol_id)

if (is.null(lego_demog_raw)) {
  stop("No data downloaded from Databrary. Are you logged in?")
}

```

### Structure of demographic data file

```{r}
str(lego_demog_raw)
```

Let's select all the variables and change the periods to underscores in the variable names.

```{r}
# Select relevant spreadsheet columns, filter those with actual session data, and rename variables to use underscores
lego_demog_raw <- lego_demog_raw %>%
  select(session_id,
         session_name,
         session_date,
         session_release,
         participant.ID,
         participant.birthdate,
         participant.gender,
         participant.race, 
         participant.ethnicity,
         participant.disability, 
         participant.language, 
         group.name,
         context.setting,
         context.language,
         context.country,
         context.state) %>%
  # filter(session_release %in% c('EXCERPTS', 'PRIVATE', 'SHARED')) %>%
  rename(participant_ID = participant.ID,
         participant_birthdate = participant.birthdate,
         participant_gender = participant.gender,
         participant_race = participant.race,
         participant_ethnicity = participant.ethnicity,
         participant_disability = participant.disability, 
         participant_language = participant.language,
         group_name = group.name,
         context_setting = context.setting,
         context_language = context.language,
         context_country = context.country,
         context.state = context.state)
```

# QA tests for each variable

## `session_name`

Are session name values consistent?
Print sorted list to detect anomalies.

```{r}
lego_demog_raw %>%
  select(., session_name) %>%
  arrange(., session_name)
```

## `session_date`

Are `session_date` values valid `YYYY-MM-DD` dates?
Use `lubridate` package to see if dates import properly.

```{r}
if (sum(is.na(lubridate::ymd(lego_demog_raw$session_date)))) {
  message("Some session_date values are invalid")
  is.na(lubridate::ymd(lego_demog_raw$session_date))
} else {
  message("All session_date values are valid.")
}
```

Are `session_date` dates reasonable?

The earliest year is `r min(lubridate::year(lego_demog_raw$session_date))`.

The most recent year is `r max(lubridate::year(lego_demog_raw$session_date))`.

## `session_release`

Are `session_release` values sensible?
That is, are there 'consistent' permissions across the sessions or missing values?

```{r}
release_sum <- xtabs(formula = ~ participant_ID + session_release, lego_demog_raw)
knitr::kable(release_sum)
```

Participants 6, 8, 14, 21, 23, 29, 30, and 35 have inconsistent `session_release` levels.

### Number of uploaded sessions?

```{r}
lego_demog_raw %>%
  group_by(., participant_ID) %>%
  summarize(., n_sessions = n()) %>%
  knitr::kable(.)
```

The following participants have less than two sessions:

```{r}
lego_demog_raw %>%
  group_by(., participant_ID) %>%
  summarize(., n_sessions = n()) %>%
  filter(n_sessions < 2) %>%
  knitr::kable(.)
```

The following participants have *more* than two sessions:

```{r}
lego_demog_raw %>%
  group_by(., participant_ID) %>%
  summarize(., n_sessions = n()) %>%
  filter(n_sessions > 2) %>%
  knitr::kable(.)
```
